---
title: "Data607 - HW3 - Regular Expressions"
author: "Vanita Hackworth"
date: "February 12, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 8 - Automated Data Collection with R
```{r}
library(stringr)
```

### Problem #3
####A)rearrange the vector so all names conform to firstname_lastname
1) Extract names data into a character vector name
```{r names-vector, eval=TRUE}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

name <- (unlist(str_extract_all(raw.data,"[[:alpha:],. ]{2,}")))
name
```
2)split name character vector into matrix of character vectors.  The pttern will be the comma in order to put those firstname into a seperate column

```{r names-split}
names_split <-str_split_fixed(name,pattern=fixed(","),n=2)
names_split
```

4) Create a new character vector consisting of all names in first name, last name format by using the str_c function to rejoin the elements on column2 to column1.  If the name was already in the proper sequence, the second column will contain spaces.  I also seerated each set of names with a double space, to make extracting easier.
```{r}
firstname_lastname <- str_c(names_split[,2],names_split[,1], sep = fixed(" "), collapse = "  ")
firstname_lastname
```
####B)Construct a logical vector indicating whether a character has a title
1) split name into a list of character vectors.  Then use str_detect to look for a period, which is in every title
```{r}
names_split
has_title <- str_detect(names_split[,1], pattern = fixed("."))
has_title
```

####B)Construct a logical vector indicating whether a character has a second name
1) use the str_detect from above to evaluate the second column for a period
```{r}
names_split
has_second_name <- str_detect(names_split[,2], pattern = fixed("."))
has_second_name
```

